Paper 2014 - A talk about Network

### 2014-08-25 anglais/paper.tex
improve paper

\begin{document}
\maketitle
\newpage

@@

\section{Introduction}
	\subsection{Computers network}

@@

	The computer networks have undergone major changes over time. Multiple projects came out and
	only some of them (a single one in fact) survived. Let's see how it works...

@@

	The mot important part of the network base is the "OSI model" (Open Systems Interconnection model) which characterizes and
	standardizes the internal functions of a communication system. All "Internet" networks use it. This model splits
	the work in 7 layers, each independant from one another. You can changer a layer without affecting the content
	of the others. It's due to this method that you can browse the web on wireless or wired connection without changing
	all your hardware.

@@

	This (single) layer enables devices to communicate together on a small area (eg: devices you have at your house,
	nothing more). The most common protocol used at this layer is Ethernet. This protocol is made to transmit data
	over the layer 1, with some collisions and errors detection. After that, to connect two devices between them, you
	need to use addresses. Ethernet use MAC addresses.
	\footnote{No, you don't need to have an "Apple Mac Computer" to have a MAC address.
	All physical ethernet network cards have got a unique one on the world.}
	\footnote{Exemple of MAC address: 00:54:77:65:78:21}
	A MAC address is 48 bits longs, that means that there are about 281,000 billion addresses possible. I think that's
	enough for your home (yes, remember, this protocol connects only a small area).

@@

	Nice, but, what exactly is IP ? IP is the protocol which enables to make the link with the underlying layer
	and global mapped address. IP uses... IP addresses.
	\footnote{Exemple of IP address: 82.212.175.9} to represents source and destination. With the fourth version of the protocol
	(commonly called IPv4), addresses are 32 bits long, which makes about 4 billions unique addresses. At the beginning of
	the "Internet time", this was absolutely huge.

	One more time, we will encapsulate this protocol on the underlying protocol. We have this schema right now:
	\begin{center}
	\includegraphics[scale=0.3]{content/layer3.eps}
	\end{center}

@@

	To understand a little bit more how that works, I'll draw an analogy with real life. Imagine a world with
	roads and a lot of buildings with flats. Roads are the layer 1 (wire), your car is the layer 2 (Ethernet), the building
	address, flat stage and flat number are described by the IP address. Now you are at your destination, you want to
	go to a specific location on this flat. When your enter the flat, you are in the lobby, you have multiple doors
	in front of you. That's the solution. Behind each doors you have one service.

@@

	This happen on layer 4 with multiple possible protocols. The most used are TCP and UDP. They are both the same purpose
	and differents ports, but work in two different ways, these ways are not important to understand how it works.
	ss
	Just keep in mind that these ports enable you to access specific services on the remote device. Like each time,
	this works with encapsulation. That's what we have after this encapsulation:
	\begin{center}
	\includegraphics[scale=0.3]{content/layer4.eps}
	\end{center}

@@

	Well, what is the layer 7 ? It's simply data you want to exchange. You can send "hello" (really just "hello"),
	and that's what the remote service (remember, the door) will receive. That's easy but to request a web page for example,
	we must provide a comprehensive way to talk with the web server. To perform that correctly, we use... one more time,
	protocols ! At this layer, we find a lot of protocols: http, ftp, ssh, irc, telnet, smtp, pop3, imap, ...

@@

	Now, our complete frame looks like that:
	\begin{center}
	\includegraphics[scale=0.3]{content/layer7.eps}
	\end{center}

@@

\section{A little deeper...}
	You have (I hope) the schema with building, flats and roads on your mind. Keep it fresh, I'll complete it to 
	understand well some devices and specific additionnal services you can find on networks.

@@

	That's all folks ! You now know basically how a computer network works. Pretty simple, right ? That's how it works
	nowadays, but computer science is a science which progresses everyday and the futur is already considered.

\section{What's next ?}
	Remember, I explained to you the layer 3 with the fourth version of "IP" (IPv4). Nowadays, IPv6 exists. In fact, the
	IPv4's 32 bits long addresses are too short now. We'll reach the limit of existing addresses very soon, that's the
	main reason for IPv6 to exists. The main new thing that IPv6 improves is that the addresses are no more on 32 bits,
	but on 128 bits. That puts the limit at approximately 3.4x$10^{38}$ addresses. This is really really big.
	This is not the only one change with IPv6, some technical details change too. IPv4 is more than 30 years old, 
	since then, we found and improved IPv6 with some missing or some deprecated features.

@@

	Finally, if you've heard of IPv8, IPv9 or something like that, forget it, they simply don't exist.
	\newpage

\section{Conclusion}
	You now know how computers networks (Internet mainly) works, with which protocols and how they communicate with each other.
	Your network card contains a MAC address used by IP to go outside your house. You have ports which specify the service
	and the application protocol which contains your requests for specific services.

@@

\section{Sources}
	\begin{itemize}
		\itemsep0em
		\item https://en.wikipedia.org/wiki/Ethernet
		\item https://en.wikipedia.org/wiki/Internet\_Protocol
		\item https://en.wikipedia.org/wiki/MAC\_address
		\item https://en.wikipedia.org/wiki/ARCnet
		\item https://en.wikipedia.org/wiki/TokenRing
		\item https://en.wikipedia.org/wiki/OSI\_model
		\item Graphics are made using https://draw.io
	\end{itemize}

>>>>

\begin{document}

\begin{titlepage}
\begin{center}
\textsc{\LARGE Haute Ecole de la Province de Liège}\\[0.5cm]
\textsc{\Large Catégorie technique}\\[4cm]

% Title
\huge \bfseries A talk about computer networks \\
\textsc{\Large by Daniel Maxime, 2326}\\[2cm]

% Author and supervisor
\textsc{\large Course: Langue Moderne}

\textsc{\large Lecturer: Véronique Randaxhe}

\vfill
{\large August 25, 2014}

\end{center}
\end{titlepage}
\newpage

@@

\section{Introduction}

	\subsection{Computers network}

@@

	The computer networks have undergone major changes over time. Multiple projects came out and
	only some of them survived. Let's see how it works...

@@

	The most important part of the network base is the "OSI model" (Open Systems Interconnection model) which characterizes and
	standardizes the internal functions of a communication system. All "Internet" networks use it. This model splits
	the work in 7 layers, each independant from one another. You can changer a layer without affecting the content
	of the others. It's due to this method that you can browse the web on wireless or wired connection without changing
	all your hardware/software.

@@

	This (single) layer enables devices to communicate together on a small area (eg: devices you have at your house,
	nothing more). The most common protocol used at this layer is Ethernet. This protocol is made to transmit data
	over the layer 1, with some errors detection. After that, to connect two devices between them, you
	need to use addresses. Ethernet use MAC addresses.
	\footnote{No, you don't need to have an "Apple Mac Computer" to have a MAC address.
	All physical ethernet network cards have got a unique one on the world.}
	A MAC address is 48 bits longs\footnote{Exemple of MAC address: 00:54:77:65:78:21},
	that means that there are about 281,000 billion addresses possible. I think that's enough for
	your home (yes, remember, this protocol connects only a small area).

@@

	Nice, but, what exactly is IP ? IP is the protocol which enables to make the link with the underlying layer
	and global mapped address. IP uses... IP addresses
	\footnote{Exemple of IP address: 82.212.175.9} to represents source and destination. With the fourth version of the protocol
	(commonly called IPv4), addresses are 32 bits long, which makes about 4 billions unique addresses. At the beginning of
	the "Internet time", this was absolutely huge.

	One more time, we will encapsulate this protocol on the underlying protocol. We have this schema right now:
	\begin{center}
	\includegraphics[scale=0.35]{content/layer3.eps}
	\end{center}

@@

	To understand a little bit more how that works, I'll draw an analogy with real life. Imagine a world with
	roads and a lot of buildings with flats. Roads are the layer 1 (wire), your car is the layer 2 (Ethernet), the building
	address, flat stage and flat number are described by the IP address. Now you are at your destination, you want to
	go to a specific location on this flat. When you enter the flat, you are in the lobby, you have multiple doors
	in front of you. That's the solution. Behind each doors you have one service.

@@

	This happen on layer 4 with multiple possible protocols. The most used are TCP and UDP. They are both the same purpose
	and differents ports, but work in two different ways, these ways are not important to understand how it works.
	Just keep in mind that these ports enable you to access specific services on the remote device. Like each time,
	this works with encapsulation. That's what we have after this encapsulation:
	\begin{center}
	\includegraphics[scale=0.35]{content/layer4.eps}
	\end{center}

@@

	Well, what is the layer 7 ? It's simply data you want to exchange. You can send "hello" (really just "hello"),
	and that's what the remote service (remember, the door) will receive. That's easy but to request a web page for example,
	we must provide a comprehensive way to talk with the web server. To perform that correctly, we use... one more time,
	protocols ! At this layer, we find a lot of protocols: http (for web pages), ftp (for file sharing),
	ssh (remote console), irc (chat), telnet (remote console), smtp (sending mail), pop3 and imap (mail storage), ...

@@

	Now, our complete frame looks like that:
	\begin{center}
	\includegraphics[scale=0.35]{content/layer7.eps}
	\end{center}

@@

\section{A little deeper...}

	You have (I hope) the schema with building, flats and roads on your mind. Keep it fresh, I'll complete it to 
	understand well some devices and specific additionnal services you can find on networks.

@@

	That's all folks ! You now know (basically) how a computer network works. Pretty simple, right ? That's how it works
	nowadays, but computer science is a science which progresses everyday and the futur is already considered.

\section{What's next ?}

	Remember, I explained to you the layer 3 with the fourth version of "IP" (IPv4). Nowadays, IPv6 exists. In fact, the
	IPv4's 32 bits long addresses are too short now. We'll reach the limit of existing addresses very soon, that's the
	main reason for IPv6 to exists. The main new thing that IPv6 improves is that the addresses are no more on 32 bits,
	but on 128 bits. That puts the limit at approximately 3.4x$10^{38}$ addresses. This is really really big.
	This is not the only one change with IPv6, some technical details change too. IPv4 is more than 30 years old, 
	since then, we found and improved IPv6 with some missing or some deprecated features.

@@

	Finally, if you've heard of IPv8, IPv9 or something like that, forget it, they simply don't exist.

\section{Conclusion}

	You now know how computers networks (Internet mainly) works, with which protocols and how they communicate with each other.
	Your network card contains a MAC address used by IP to go outside your house. You have ports which specify the service
	and the application protocol which contains your requests for specific services.

@@

\section{Glossary}

	\begin{itemize}
		\itemsep0em
		\item \textbf{Protocols}:
			specific method used to describe the mechanism, the fields of data, bits, bytes,
			the order of operations and data to send and receive to make a working-communication.

		\item \textbf{Internet Provider}: 
			companies interconnected with other companies which enables you to reach different networks.
			All of these different networks make "the Internet", it's simply a lot of interconnected-networks.

		\item \textbf{Checksum}: 
			a fast operation which consist of make the sum of the data with an algorythm more or less secure.
			This mechanism enables protocol to know if the payload is altered or not by doing the same operation
			and comparing the checksum value. Safer is the checksum, longer and slower it is.

		\item \textbf{Data encapsulation}: 
			the fact of put data on a larger container.

		\item \textbf{Response mechanism}: 
			when you receive data, you can response to it just by fill in the application payload
			with your response, then swaping source and destination address on headers.

		\item \textbf{Subnet}: 
			each IP contained on the same subnets can communicate each other without router, a router
			is the device that enables communication between multiple subnets.

		\item \textbf{TCP}: 
			protocol which enables "secure" connection with a negociation (handshake) before working. It provides
			an acknowledgment and retransmission light but optimized system.

		\item \textbf{UDP}: 
			protocol less "secure" than TCP because it does not support retransmission, acknowledgment and handshake,
			but is very smaller than TCP. This protocol is often used for multimedia content streaming.

		\item \textbf{Switch}: 
			devices that works on layer 2 only and forward frame on the right (destination address) physical port.

		\item \textbf{Router}: 
			interconnect multiple subnet and take decision for routing (best path, etc.)

		\item \textbf{Firewall}: 
			security software which analyze traffic and take decision based on users rules.

	\end{itemize}

\section{Sources}
	\begin{itemize}
		\itemsep0em
		\item https://en.wikipedia.org/wiki/Ethernet
		\item https://en.wikipedia.org/wiki/Internet\_Protocol
		\item https://en.wikipedia.org/wiki/MAC\_address
		\item https://en.wikipedia.org/wiki/ARCnet
		\item https://en.wikipedia.org/wiki/TokenRing
		\item https://en.wikipedia.org/wiki/OSI\_model
		\item http://www.cisco.com/c/en/us/support/docs/ip/routing-information-protocol-rip/13769-5.html
		\item Graphics are made using https://draw.io
	\end{itemize}
